Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2020-01-17 IEEE_ConferenceOnGames/wcci_2020.tex
Revised introduction

\begin{abstract}
In this paper, we propose an evolutionary algorithm that follows an open-ended
evolution approach to generate levels for the Angry Birds video game. The levels
themselves are composed of a set of birds that the player can throw with a
slingshot, a certain amount of pigs that they must destroy, and a given number
of pieces that conform structures that may or may not protect the pigs from
birds thrown at them. The current goal is the generation of diverse structures
that are playable in the game, having the additional characteristics of being
fun and enjoyable. We propose a multi-layered search, first constructing
composite structures from basic blocks, to then build more complex structures
building from these composites. We follow an open-ended evolution approach in
which the evolution of structures is not guided towards a single objective but
is rather free to evolve and generate novelty or diversity. The fitness function
we use to evaluate the proposed levels considers how complex levels have become
and how different they are from the rest of the population. The experiments
conducted show that an evolutionary approach allows the generation of levels
that are novel and interesting to play.
% Another revision to English would be good.
% Done - Mario
\end{abstract}

\begin{IEEEkeywords}
Genetic algorithm, procedural generated content, open-ended evolution,
evolutionary computation
\end{IEEEkeywords}

\section{Introduction}
The video game industry is always looking for content creators to satisfy
the demand for high-quality content. Current development teams include game
designers,  artists, animators,  writers, level designers, among others.
Procedural Content Generation (PCG) is often applied to generate many
aspects of the game, from maps, levels or characters, to narrative and game
design. PCG techniques can lower the cost in both time and human resources
when generating content for different video game projects, without
compromising quality \cite{Yannakakis2017,YannakakisContentGeneration}. In
this paper, we introduce an evolutionary content generation system, to
generate fun and exciting levels for the popular Angry Birds video game. We
follow an Open-Ended Evolution (OEE) approach.  OOE is a type of genetic
evolution algorithm with a different objective. In this case, the goal is
not to reach a final state of in the evolution, for instance, to reach an
optimal solution in a search space. An OEE aims to continue evolving and
creating new types of entities, perhaps more complex or novel on each new
generation. Entities evolve not only inside the frame of their species, but
they can also create entirely new species \cite{Standish2003}. Taylor et al.
\cite{Taylor2016, Taylor} define OEE as a system capable of creating novelty
beyond a point where nothing changes on the current group of interest rather
than just converging to a stable or quasi-stable state. For this project,
OEE is defined simply as a genetic algorithm with different classes of
objects, with compositional and taxonomic relationships between them.
Objects in a class can evolve in parallel or as in this case, independently
of each other. The object-oriented model follows these principles to model
complex behavior and structure \cite{levesque1986knowledge}.

The case study for this paper is to follow the rules of a level generation
competition for the Angry Birds video game developed by Rovio Entertainment.
We briefly explain the game.  The objective is to destroy structures placed
on a map as well as eliminating green pig-like characters. In order to
obtain the highest score possible, the game uses a gravity and collision
system that allows the structures to fall and be destroyed by means of
damage. The structures have  three different construction materials
\textit{ice}, \textit{wood} and \textit{stone}, each one with a different
resistance, ranging from fragile, normal and sturdy in that order. The game
mechanic is to use a slingshot to shoot a certain number of birds with
different abilities, to destroy both the structures and the objectives. A
high score is achieved by causing the most destruction throwing the least
birds possible \cite{RovioEntertainmentCorporation2009}. The main objective
of this work is to generate levels that can be played by naïve or
intelligent agents as well as human players. The levels themselves need to
be complex in the sense that interesting structures can be created, but at
the same time, remaining possible to complete in some way \cite{Stephenson,
Stephenson2018}.
% What is the problem?

One peculiarity of PCG is the richness of problems and domains in which it can
be applied.  This is because each type of game dictates different restrictions
and mechanics. In this case, we have additional restrictions, dictated by the
competition track \cite{Renz}. This case study provides an interesting challenge
for the participating algorithms. The proposed PCG algorithm has to be flexible
on the pieces it can use, and also consider the running time, as both are
determined at runtime. The aim is to develop a software capable of creating game
levels having the objective of being both \textit{interesting} and
\textit{playable}. Candidate solutions, in this case, Levels of the game, will
be evaluated by using the open-source software Science Birds \cite{sciencebirds}
which uses the same resources used on the original angry birds game. Using this
software for the evaluation is an additional restriction because we are limited
to the data collected during the simulation.  
% What is learned?
In this work, we found that an evolutionary algorithm can be used to create this
type of content, because it is not only able to create levels that are
interesting and diverse but also capable of producing levels an average human
could have trouble imagining. The rest of the paper is as follows. Section
\ref{SoA}, contains a quick overview of some aspects used by different authors
to tackle this problem, their design, and basic functionality. In Section
\ref{prop_method} we describe our current proposal.  How we modeled the
candidate solutions and the way we defined our fitness function to determine the
playability and complexity of the generated levels. Finally, in Section
\ref{conclusion}, we provide our conclusions as well as improvements that can
help to refine the proposed work.

>>>>

\begin{abstract}
In this paper, we propose an evolutionary algorithm that follows an open-ended
evolution approach to generate levels for the Angry Birds video game. The levels
themselves are composed of a set of birds that the player can throw with a
slingshot, a certain amount of pigs that they must destroy, and a given number
of pieces that conform structures that may or may not protect the pigs from
birds thrown at them. The current goal is the generation of diverse structures
that are playable in the game, having the additional characteristics of being
fun and enjoyable. We propose a multi-layered search, first constructing
composite structures from basic blocks, to then build more complex structures
building from these composites. We follow an open-ended evolution approach in
which the evolution of structures is not guided towards a single objective but
is rather free to evolve and generate novelty or diversity. The fitness function
we use to evaluate the proposed levels considers how complex levels have become
and how different they are from the rest of the population. The experiments
conducted show that an evolutionary approach allows the generation of levels
that are novel and interesting to play.
\end{abstract}

\begin{IEEEkeywords}
Genetic algorithm, procedural generated content, open-ended evolution,
evolutionary computation, games.
\end{IEEEkeywords}

\section{Introduction}

The video game industry is always looking for content creators in order to satisfy
the demand for high-quality, and diverse, content. Current development teams include game 
designers,  artists, animators,  writers, and level designers, among others.
Procedural Content Generation (PCG) is often applied to generate many
aspects of the game, from maps, levels or characters, to narrative and overall
design. PCG techniques can lower the cost in both time and human resources
when generating content for different video game projects, without
compromising quality \cite{Yannakakis2017,YannakakisContentGeneration}. In
this paper, we introduce an evolutionary content generation system that is able to 
generate fun and exciting levels for the popular Angry Birds video game \cite{RovioEntertainmentCorporation2009}.

We are goint to follow an Open-Ended Evolution (OEE) approach.  OEE is a type of genetic evolution algorithm in which the goal is
not so much to reach a final, possible optimal, state of in the evolution, for instance but to continue evolving and
creating new types of entities, perhaps more complex or novel on each new
generation. Entities evolve not only inside the context of their species, but
they can also create entirely new species \cite{Standish2003}. Taylor et al.
\cite{Taylor2016, Taylor} define OEE as a system capable of creating novelty
beyond a point where nothing changes on the current group of interest rather
than just converging to a stable or quasi-stable state. For this project,
OEE is defined simply as a genetic algorithm with different classes of
objects, with compositional and taxonomic relationships between them.
Objects in a class can evolve in parallel or as in this case, independently
of each other. The object-oriented model follows these principles to model
complex behavior and structure \cite{levesque1986knowledge}. % Maybe specify what you mean by "classes" in this context - JJ

The case study for this paper is to follow the rules of a level generation
competition for the Angry Birds video game developed by Rovio Entertainment.
We briefly explain the game. % Already mentioned above. Maybe it would be better to move this paragraph to where you mention it for the first time - JJ
The objective is to destroy structures placed
on a map as well as eliminating green pig-like characters. In order to
obtain the highest score possible, the game uses a gravity and collision
system that allows the structures to fall and be destroyed by means of
damage. The structures have  three different construction materials
\textit{ice}, \textit{wood} and \textit{stone}, each one with a different
resistance: fragile, normal or sturdy. The game
mechanic is to use a slingshot to shoot a certain number of birds with
different abilities, to destroy both the structures and the objectives. A
high score is achieved by causing the most destruction throwing the least
birds possible \cite{RovioEntertainmentCorporation2009}. The main objective
of this work is to generate levels that can be played by naïve or
intelligent agents as well as human players. The levels themselves need to
be complex in the sense that interesting structures can be created, but at
the same time, remaining possible to complete in some way \cite{Stephenson,
Stephenson2018}.
% What is the problem? Please formulate this with more precision - JJ

One peculiarity of PCG is the richness of problems and domains in which it can
be applied.  This is because each type of game dictates different restrictions
and mechanics. In this case, we have additional restrictions, dictated by the
competition track \cite{Renz}. % make some mention to why this is important - JJ
This case study provides an interesting challenge
for the participating algorithms. The proposed PCG algorithm has to be flexible
on the pieces it can use, and also consider the running time, as both are
determined at runtime. The aim is to develop a software capable of creating game
levels having the objective of being both \textit{interesting} and
\textit{playable}. Candidate solutions, in this case, Levels of the game, will
be evaluated by using the open-source software Science Birds \cite{sciencebirds}
which uses the same resources used on the original angry birds game. Using this
software for the evaluation is an additional restriction because we are limited
to the data collected during the simulation.  
% What is learned?
In this work, we found that an evolutionary algorithm can be used to create this
type of content, because it is not only able to create levels that are
interesting and diverse but also capable of producing levels an average human
could have trouble imagining. The rest of the paper is as follows. Section
\ref{SoA}, contains a quick overview of some aspects used by different authors
to tackle this problem, their design, and basic functionality. In Section
\ref{prop_method} we describe our current proposal.  How we modeled the
candidate solutions and the way we defined our fitness function to determine the
playability and complexity of the generated levels. Finally, in Section
\ref{conclusion}, we provide our conclusions as well as improvements that can
help to refine the proposed work.
